
#7
##a
$P(Y=y) = \frac{\lambda^y}{y!}e^{-\lambda},y=0,1,…$
For Y_1;Y_2; : : : ;Y_n iid Poisson random variables will have a joint frequency function that is
a product of the marginal frequency functions, the log likelihood will thus be:
$l(\lambda)=\sum_{i=1}^n(Y_ilog\lambda-\lambda-logY_i!)$
$=log\lambda\sum_{i=1}^nY_i-n\lambda-\sum_{i=1}^nlogY_i!$
We need to find the maximum by finding the derivative: 
$l'(\lambda) = \frac{1}{\lambda}\sum_{i=1}^ny_i-n=0$
Which implies that the estimate should be
      $\hat{\lambda}=\overline{y}$
      as Poisson distrabution we know
      $\lambda=\hat{\mu}$
      then
      $\hat{\mu}=\overline{y}$
      
##b
```{r}
set.seed(1111)
sample <- rpois(30,10)
mean(sample)
var(sample)
sum(sample)
```


```{r}
#gamma(x1+1)…gamma(xn+1)
ga<-function(x){
ga<-1
for(i in 1:length(x)){
ga<-ga*gamma(x[i]+1)
}
ga
}
```


```{r}
hist(sample)
```


```{r}
#gamma(x1+1)…gamma(xn+1)
ga(sample)
```


```{r}
f<-function(lambda){f = (lambda)-30*lambda+309*log(lambda)-log(4.315389*10^211)
return(-f)
  }
```


```{r}
samp <- nlminb(start=10,objective=f,lower=8,upper=12)
```


```{r}
mu_hat<-samp$par
as.vector(sample)

```
```{r}
x=seq(6,18,by = 2)

```
```{r}
#HIstogram & the redline is $\mu_hat$
hist(sample,freq = F)
abline(v = mu_hat,col = 'red')
```
##c
```{r}
plot(sample)
```


###1
```{r}
down1 = as.numeric(-1.96*(mu_hat/30)^(1/2)+mu_hat)
up1 = as.numeric(1.96*(mu_hat/30)^(1/2)+mu_hat)

```
$[-1.96\sqrt{Var(\hat{\mu})/n}+\hat\mu,1.96\sqrt{Var(\hat{\mu})/n}+\hat\mu]=$[down1,up1]
###2
```{r}
down2 = as.numeric(-1.96*(mu_hat/30)^(1/2)+mu_hat)
up2 = as.numeric(1.96*(mu_hat/30)^(1/2)+mu_hat)
```
$[\hat{\mu} - 1.96\sqrt{\hat{\mu} / n},\hat{\mu} + 1.96\sqrt{\hat{\mu} / n}]$=[down2,up2]
###3
```{r}
B=5000
booty = lapply(1:B, function(i) sample(sample, replace = T))
boot.mean = sapply(booty,mean)
boot.var = sapply(booty,var)
boot.t = (boot.mean - mu_hat)/sqrt(boot.var/30)
a=quantile(boot.t,probs=c(0.025,0.975))

down3=mu_hat+as.numeric(a[1])
up3=mu_hat+as.numeric(a[2])

```
#plot
```{r}
plot(sample,main = 'Scatterplot',lab=c(10,12,12) )
down3
abline(h=c(down1,up1,down2,up2,down3,up3),col = c('red','red','blue','blue','green','green'))

```


```{r}
library(ggplot2)
sp <- as.data.frame(sample)
sp$index <- factor(row.names(sp),levels = c(1:30))
sp= sp[,c(2,1)]

qplot(index,sample,data = sp,ylab = 'number',main = 'scatterplot')+geom_hline(yintercept = c(down1,up1,down2,up2,down3,up3),col = c('red','red','blue','blue','green','green'))+annotate(geom = "text", x = 20, y = 20, hjust = 0,
    label="95%CI q1&q2-beteween the blue lines
    q3 beteween the green lines","rect", ymin=down1, ymax=up1, alpha=.1,fill="blue")
```


